---
title: "Multiple Regression on Social Media Data"
author: "Gauri Raghuram"
output: html_document
---

```{r}
library(factoextra)
library(FactoMineR)
library(psych)
```


### **Predicting the Impact on Mood Productivity based on the amount of Social Media Used**

```{r}
social_media <- read.csv("D:/Rutgers/R2nd Semester/Multivariate/Socialmediacleaned.csv",row.names=1)

socialmedia2 <- social_media[, c("Instagram_Hours",	"LinkedIn_Hours",	"Snapchat_Hours", "Twitter_Hours",	"Whatsapp_Wechat_hours", "Reddit_hours",	"Youtube_hours",	"OTT_hours", "Mood_Productivity")]


```

```{r}
str(socialmedia2)
```

* It cab be observed that the 'Mood_Productivity' column is of 'chr' (Character) data type which as two unique values: "Yes" and "No".
* To conduct Multiple Regression, we would need to convert Mood Productivity into a numeric column with values 0 and 1.


**Converting the Character column into Numeric column**

```{r}

socialmedia2$Mood_Productivity_num <- ifelse(socialmedia2$Mood_Productivity == "Yes", 1, 0)

str(socialmedia2)
```



**Conducting Multiple Regression - Using the mood_productivity as the target variable**

```{r}
# Performing multiple regression on Social Media dataset
fit <- lm(Mood_Productivity_num~Instagram_Hours+LinkedIn_Hours+Snapchat_Hours+Twitter_Hours+Whatsapp_Wechat_hours+Reddit_hours+Youtube_hours+OTT_hours, data=socialmedia2)

#show the results
summary(fit)


```

**Inference:**

* The Median being close to 0 showcases that the model is able to predict perfectly.
* Based on the residual data provided, it can be inferred that residuals may be following a Normal Distribution.
* Furthermore, it can be observed that most columns have the P-values > 0.05, which means that none of them are significantly affecting the target variable.
* However, the F-Statistic value is just above 1 which means it is a good model but not a great model.


#### **Trying to fit the model by removing 'Snapchat' to see if makes any difference in the model summary**

```{r}
# Performing multiple regression on Social Media dataset
fit_1 <- lm(Mood_Productivity_num~Instagram_Hours+LinkedIn_Hours+Twitter_Hours+Whatsapp_Wechat_hours+Reddit_hours+Youtube_hours+OTT_hours, data=socialmedia2)

#show the results
summary(fit_1)
```

**Inference:**

* By removing the 'Snapchat' column, it can be inferred that the model is still performing below expectations.
* Therefore, analysis using the previous fit model can be continued.


```{r}
coefficients(fit)
```

#### **Residual Analysis**

```{r}
library(GGally)
ggpairs(data=socialmedia2, title="Social Media Data")
```

**Inference:**

* The graph clearly indicates that none of the columns have a linear relationship.

#### **Confidence Level**

```{r}
confint(fit,level=0.95)
```

**Inference:**

* Gives a range (low - high)

#### **Predictions**

```{r}
fitted(fit)
```

**Inference:**

* Here, the model is predicting mood productivity for every student.


#### **Residual Fit**

```{r}
residuals(fit)
```

**Inference:**

* Ideally, the errors should always be close to 0.
* All of the values in our case also seem to be close to 0.
* It gives the difference between the actual value and the predicted value.

#### **Residual Chart**

```{r}
plot(fit)
```

**Inference:**

* The plots above are heavily relied upon to verify whether the model's assumptions have been met. They help in identifying significant model issues.
* And we can also see that the line is not that far from the predicted line.


### Model Accuracy

```{r}
summary(fit)$r.squared
```

### Inference
* The accuracy is 45%.

#### **As observed the model may not have been performing well. Therefore, let's conduct PCA and carry out Multiple Regression on the PCs to check if that provides any improvement**

```{r}
social_media_num <- socialmedia2[, c("Instagram_Hours",	"LinkedIn_Hours",	"Snapchat_Hours", "Twitter_Hours",	"Whatsapp_Wechat_hours", "Reddit_hours",	"Youtube_hours",	"OTT_hours", "Mood_Productivity_num")]

social_media_num_pca <- socialmedia2[, c("Instagram_Hours",	"LinkedIn_Hours",	"Snapchat_Hours", "Twitter_Hours",	"Whatsapp_Wechat_hours", "Reddit_hours",	"Youtube_hours",	"OTT_hours")]

social_media_pca <- prcomp(social_media_num_pca, scale. = TRUE)

social_media_pca

```

```{r}
summary(social_media_pca)
```


**Scree Plot**

**To further understand the ideal number of PCs, we can carry out the Scree Plot**

```{r}
(eigen_social_media <- social_media_pca$sdev^2)
names(eigen_social_media) <- paste("PC",1:8,sep="")

plot(eigen_social_media, xlab = "Component number", ylab = "Component variance", type = "l", main = "Scree diagram")
```

**Inference**:

* Based on the observed Scree plots, **3 can be chosen as the ideal number of Principal Components**


#### **Multiple Regression using 3 PCs**

```{r}

PC1 <- social_media_pca$x[, 1]
PC2 <- social_media_pca$x[, 2]
PC3 <- social_media_pca$x[, 3]

# Create a new dataset with selected principal components
pc_data <- data.frame(PC1, PC2, PC3)

#Combining Target Variable to pc_data
pc_data$mood_prod_num <- socialmedia2$Mood_Productivity_num

```

```{r}
pc_data

```


```{r}
fit_pc <- lm(mood_prod_num~PC1+PC2+PC3, data=pc_data)

summary(fit_pc)

```

**Inference:**

* Here, all the PCs can be seen to have a high P-value (>0.05).
* The next best method to deal with this could be to carry out **Factor Analysis** to check for a better model.


#### **Factor Analysis and Multiple Regression on FCs**

**Scree Plot -- To check for the ideal number of factors**

```{r}
fa.parallel(social_media_num_pca) 
```

**Inference**:

* Following the FA Actual Data, we can interpret as **2 to be the ideal number of factors**.


**Factor Model**

```{r}
fit.pc <- principal(social_media_num_pca, nfactors=2, rotate="varimax")

round(fit.pc$values, 3)
fit.pc$loadings

```

#### **Multiple Regression using 2 RCs**

```{r}
loadings <- fit.pc$scores[, c("RC1", "RC2")]
loadings_df <- as.data.frame(loadings)
loadings_df <- round(loadings_df, 3)
```

```{r}
loadings_df$mood_prod_num <- socialmedia2$Mood_Productivity_num
```

```{r}
fit_rc <- lm(mood_prod_num~RC1+RC2, data=loadings_df)

summary(fit_rc)
```

**Inference:**

* Here, all the RCs can be seen to have a high P-value (>0.05) indicating that this may not be the right course of action.


#### **Summary**

* The overall analysis has made us understand how multiple regression is performing poorly on columns. 
* Furthermore, even after performing PCA and Factor Analysis we saw the the multiple regression is still performing poorly.


