---
title: "Logistic Regression Apple's Quality"
output: html_document
date: "2024-04-19"
---


```{r}
library(factoextra)
library(FactoMineR)
library(psych)
library(ggplot2)
library(cowplot)
library(caret)
library(e1071)
library(pROC)
```

### Model Development

```{r}
apples <- read.csv("D:/Rutgers/R2nd Semester/Multivariate/apple2.csv",row.names=1)
apples
apples2 <- apples[, c("Size", "Weight", "Crunchiness", "Sweetness", "Juiciness", "Ripeness", "Acidity","Quality")]
```

```{r}
head(apples2)
```

### Performed the following to check the data type of the columns

```{r}
str(apples2)
```

### Converting Quality to integer since its a charater column.

```{r}
apples2$Quality_num<- ifelse(apples2$Quality == "good", 1, 0)

str(apples2)
```

### Converting into factors

```{r}
apples2[apples2 == "?"] <- NA
apples2$Quality <- ifelse(test=apples2$Quality =="good", yes="good", no="bad") 
apples2$Quality <- as.factor(apples2$Quality)


```

### Inference
* We are converting the categorical column, which is our predictor column into factors to make it easier for R to recognize it as a categorical column .
* We have converted it into factors with 2 levels.




```{r}
str(apples2)
```

### Inference
* We can see that Quality column has changed to factors with 2 levels.

### Performing Logistic Regression model

```{r}
logistic <- glm(Quality ~ ., data=apples2, family="binomial")
summary(logistic)
```

### Inference
* As we can observe the p value is 1 for all the outcome variables, which means that there is a weak relationship between the predictor and outcome variable.

### Residual Analysis

```{r}
predicted.data1 <- data.frame(probability.of.Quality=logistic$fitted.values,Quality=apples2$Quality)
predicted.data1 <- predicted.data1[order(predicted.data1$probability.of.Quality, decreasing=FALSE),]
predicted.data1$rank <- 1:nrow(predicted.data1)
ggplot(data=predicted.data1, aes(x=rank, y=probability.of.Quality)) +
geom_point(aes(color=Quality), alpha=1, shape=4, stroke=2) +
xlab("Index") +
ylab("Predicted probability of Quality")
```

### Inference
* We can observe that there is not much overlap in the residual analysis model performed.
* This shows the independence of variables, absence of outliers.
* If there is no overlap we can tell that assumptions of logistic regression are met.

### Prediction

```{r}
pdata1 <- predict(logistic,newdata=apples2,type="response" )
pdata1
```

### Inference
* Here the model is predicting Quality for every apple.

### Confusion matrix with model accuracy

```{r}
pdataF1 <- as.factor(ifelse(test=as.numeric(pdata1>0.5) == "good", yes="good", no="bad"))
pdataF1 <- factor(pdataF1, levels = levels(apples2$Quality))

confusionMatrix(pdataF1, apples2$Quality)
```

### Inference
* We can see the models overall performance if it is correctly classifying into its related classes or not.
* Also tells us about sensitivity, specificity etc.
* The model accuracy is also shown which is 0.49, where as balanced accuracy is 0.50.
