---
title: "Mean and Variance Analysis"
output: html_document
date: "2024-02-24"
---

#### Firstly, I have imported my dataset to check the columns, its datatype and other related properties.

```{r}
library(readr)
apples <- read_csv("D:/Rutgers/R2nd Semester/Multivariate/apple_quality.csv")
attach(apples)
str(apples)
```

#### The below code gives the values of the columns as per the dataset.

```{r}
apples$Quality <- as.factor(apples$Quality)
apples_x <- apples[, 2:8]
apples_x
```

#### The below code is used to check the mean of the columns, the covariance between them and vector (apples_d) where each element corresponds to the distance of the corresponding observation from the mean, considering the covariance structure.


```{r}
apples_cm <- colMeans(apples_x)
apples_S <- cov(apples_x)
apples_d <- apply(apples_x, MARGIN = 1, function(apples_x)t(apples_x - apples_cm) %*% solve(apples_S) %*% (apples_x - apples_cm))
apples_cm
apples_S
apples_d 
```

### T-Tests
#### I wanted to perform t-tests on each of the characteristic of the apple to check which set of apples which had a particular characteristic had more tendency to have good quality or bad quality. Wanted to check the significance.

```{r}
with(data=apples,t.test(Size[Quality=="good"],Size[Quality=="bad"],var.equal=TRUE))
```

### Inference 
#### It can be observed that Size is significantly affecting the quality of the apple because its p value is less than 0.05. 

```{r}
with(data=apples,t.test(Weight[Quality=="good"],Weight[Quality=="bad"],var.equal=TRUE))## Weight is not significantly affecting the quality of the apples.
```

### Inference
#### It can be observed that Weight is not significantly affecting the quality of the apples because the p value is greater than 0.05.

```{r}
with(data=apples,t.test(Sweetness[Quality=="good"],Sweetness[Quality=="bad"],var.equal=TRUE))
```

### Inference
#### It can be observed that Sweetness is significantly affecting the quality of the apple because its p value id too less than 0.05.

```{r}
with(data=apples,t.test(Crunchiness[Quality=="good"],Crunchiness[Quality=="bad"],var.equal=TRUE))
```

### Inference
#### It can be observed that Crunchiness is not significantly affecting the quality of the apples because the p value is greater than 0.05.

```{r}
with(data=apples,t.test(Juiciness[Quality=="good"],Juiciness[Quality=="bad"],var.equal=TRUE))
```

### Inference
#### It can be observed that Juiciness is significantly affecting the quality of the apple because its p value id too less than 0.05.

```{r}
with(data=apples,t.test(Ripeness[Quality=="good"],Ripeness[Quality=="bad"],var.equal=TRUE))
```

### Inference
#### It can be observed that Ripeness is significantly affecting the quality of the apple because its p value id too less than 0.05.

```{r}
with(data=apples,t.test(Acidity[Quality=="good"],Acidity[Quality=="bad"],var.equal=TRUE))
```

### Inference
#### It can be observed that Acidity is significantly affecting the quality of the apple because its p value id too less than 0.05.

#### After performing t-test on all of the characteristics, I observed the p values are very low for most of the characteristics, hence I did not scale the values down and perform the t-tests again.


### Multivariate Analysis
#### I wanted to perform multivariate analysis and check the p value when all the characteristics are considered together.

```{r}
library(Hotelling)
t2testapple <- hotelling.test(Size+ Weight + Sweetness + Crunchiness + Ripeness + Acidity ~ Quality, data=apples)

cat("T2 statistic =",t2testapple$stat[[1]],"\n")
print(t2testapple)
```

### Inference
#### It can be observed that the p value is 0, which is telling that the since the p values of most of the characteristic is 0 except for three of them, they very strongly significantly affecting the quality of the apple. 


#### I wanted to check the p value when I considered the characterictics which gave me p values greater than 0.05.

```{r}
library(Hotelling)
t2testapple <- hotelling.test(Weight + Crunchiness + Acidity ~ Quality, data=apples)

cat("T2 statistic =",t2testapple$stat[[1]],"\n")
print(t2testapple)
```

### Inference
#### It can be observed that the p value is greater than 0.05, which tells that the weight, Crunchiness and Acidity are not significatly affecting the quality of the apple.




### Levene Test
#### I wanted to check with how much significance each characteristic of the apple is affecting its quality using levene test.

```{r}
library(car)
leveneTest(Size ~ Quality, data=apples)
```

### Inference
#### It can be observed that the size of the apple is significantly affecting the quality of the apple.

```{r}
leveneTest(Weight ~ Quality, data=apples)

```

### Inference
#### It can be observed that the weight of the apple is significantly affecting the quality of the apple.

```{r}
leveneTest(Sweetness ~ Quality, data=apples)

```

### Inference
#### It can be observed that the sweetness of the apple is not that significantly affecting the quality of the apple.

```{r}
leveneTest(Crunchiness ~ Quality, data=apples)

```

### Inference
#### It can be observed that the crunchiness of the apple is significantly affecting the quality of the apple.

```{r}
leveneTest(Juiciness ~ Quality, data=apples)

```

### Inference
#### It can be observed that the Juiciness of the apple is significantly affecting the quality of the apple.

```{r}
leveneTest(Ripeness ~ Quality, data=apples)

```
### Inference
#### It can be observed that the Ripeness of the apple is a little less significantly affecting the quality of the apple.

```{r}
leveneTest(Acidity ~ Quality, data=apples)
```

### Inference
#### It can be observed that the Acidity of the apple is significantly affecting the quality of the apple.


### Analysis of Variance Test
#### I wanted to check with how much significance each characteristic of the apple is affecting its quality using analysis of varience test.

```{r}
summary(aov(Size ~ Quality))

```

### Inference
#### It can be observed that the size of the apple is significantly affecting the quality of the apple.


```{r}
summary(aov(Weight ~ Quality))

```

### Inference
#### It can be observed that the weight of the apple is not very significantly affecting the quality of the apple.

```{r}
summary(aov(Sweetness ~ Quality))

```

### Inference
#### It can be observed that the sweetness of the apple is significantly affecting the quality of the apple.

```{r}
summary(aov(Crunchiness ~ Quality))

```

### Inference
#### It can be observed that the crunchiness of the apple is not very significantly affecting the quality of the apple.

```{r}
summary(aov(Juiciness ~ Quality))

```

### Inference
#### It can be observed that the Juiciness of the apple is significantly affecting the quality of the apple.

```{r}
summary(aov(Ripeness ~ Quality))

```

### Inference
#### It can be observed that the Ripeness of the apple is significantly affecting the quality of the apple.

```{r}
summary(aov(Acidity ~ Quality))
```

### Inference
#### It can be observed that the Acidity of the apple is not very significantly affecting the quality of the apple.


### F-Tests
#### I wanted to check with how much significance each characteristic of the apple is affecting its quality using F test.

```{r}
attach(apples)
var.test(Size[Quality=="good"],Size[Quality=="bad"])
```

### Inference
#### It can be observed that the Size of the apple is significantly affecting the quality of the apple.

```{r}
attach(apples)
var.test(Weight[Quality=="good"],Weight[Quality=="bad"])
```

### Inference
#### It can be observed that the weight of the apple is significantly affecting the quality of the apple.

```{r}
attach(apples)
var.test(Sweetness[Quality=="good"],Sweetness[Quality=="bad"])
```

### Inference
#### It can be observed that the Sweetness of the apple is not significantly affecting the quality of the apple.

```{r}
attach(apples)
var.test(Crunchiness[Quality=="good"],Crunchiness[Quality=="bad"])
```

### Inference
#### It can be observed that the crunchiness of the apple is significantly affecting the quality of the apple.

```{r}
attach(apples)
var.test(Juiciness[Quality=="good"],Juiciness[Quality=="bad"])
```

### Inference
#### It can be observed that the Juiciness of the apple is significantly affecting the quality of the apple.


```{r}
attach(apples)
var.test(Ripeness[Quality=="good"],Ripeness[Quality=="bad"])
```

### Inference
#### It can be observed that the Ripeness of the apple is significantly affecting the quality of the apple.

```{r}
attach(apples)
var.test(Acidity[Quality=="good"],Acidity[Quality=="bad"])
```

### Inference
#### It can be observed that the Acidity of the apple is significantly affecting the quality of the apple.
