---
title: "Multiple Regression on apples dataset"
output: html_document
date: "2024-04-14"
---

```{r}
library(factoextra)
library(FactoMineR)
library(psych)
```


### Predicting the Impact on Quality of apple based on the different characteristics of apple

```{r}
apples <- read.csv("D:/Rutgers/R2nd Semester/Multivariate/apple2.csv",row.names=1)
apples
apples1 <- apples[, c("Size", "Weight", "Crunchiness", "Sweetness", "Juiciness", "Ripeness", "Acidity","Quality")]
```

```{r}
str(apples1)
```

### Inference
* It cab be observed that the 'Quality' column is of 'chr' (Character) data type which as two unique values: "good" and "bad".
* To conduct Multiple Regression, we would need to convert Quality into a numeric column with values 0 and 1.

### Converting the Character column into Numeric column

```{r}
apples1$Quality_num<- ifelse(apples1$Quality == "good", 1, 0)

str(apples1)
```

### Conducting Multiple regressions keeping the quality as the target variable

```{r}
# Performing multiple regression on Social Media dataset
fit_apple <- lm(Quality_num~Size+Weight+Crunchiness+Sweetness+Juiciness+Ripeness+Acidity, data=apples1)

#show the results
summary(fit_apple)
```

### Inference
* The Median being close to 0 showcases that the model is able to predict perfectly.
* Based on the residual data provided, it can be inferred that residuals may be following a Normal Distribution.
* Furthermore, it can be observed that Sweetness, Size are the one's which are significantly affecting the target variable.
* However, the F-Statistic value is above 1 which means it is a good model but not a great model.


### Trying to fit the model by removing 'Ripeness' to see if makes any difference in the model summary

```{r}
# Performing multiple regression on Social Media dataset
fit1_apple <- lm(Quality_num~Size+Weight+Crunchiness+Sweetness+Juiciness+Acidity, data=apples1)

#show the results
summary(fit1_apple)
```

### Inference
* We can see that there is not much difference compared to the first model which we performed.




```{r}
coefficients(fit_apple)
```

### Residual Analysis

```{r}
library(GGally)
ggpairs(data=apples1, title="Quality of Apples")
```

### Inference
* The graph clearly indicates that none of the columns have a linear relationship.


### Confidence Levels

```{r}
confint(fit_apple,level=0.95)
```

### Inference
* Gives a range (low - high)

### Predictions

```{r}
fitted(fit_apple)
```

### Inference
* Here, the model is predicting mood productivity for every apple.

### Residual Fits

```{r}
residuals(fit_apple)
```

### Inference

* Ideally, the errors should always be close to 0.
* All of the values in our case also seem to be close to 0.
* It gives the difference between the actual value and the predicted value.

### Residual chart

```{r}
plot(fit_apple)
```

### Inference
* The plots above are heavily relied upon to verify whether the model's assumptions have been met. They help in identifying significant model issues.
* And we can also see that the line is far from the predicted line.


### Model Accuracy

```{r}
summary(fit_apple)$r.squared
```

### Inference
* The accuracy is about 48%.
